\textbf{\mytitle}

### 2025-06-01 draft.tex
Fix

\begin{table}[h]
    \caption{Summary table of persons involved in the project.}
    \centering
    \resizebox{\textwidth}{!}{%
\begin{tabular}{|c|l|c|c|l|}
    \hline
    Last, First name & Affiliation & Involvement (h/week) & Duration (months) & Role  \\\hline\hline
    \textbf{Orrù},  Michele & CNRS & 10 & 10 (Jul-Apr) & Project Coordinator \\
    \textbf{Chiesa}, Alessandro & EPFL & 5 & 10 (Jul-Apr) & Technical Advisor \\
    \textbf{TBD} & Ethereum Foundation & 20 & 10 (Jul-Apr) & PhD Student (Lean expert) \\
    \textbf{TBD} & Ethereum Foundation & 20 & 10 (Jul-Apr) & PhD Student (Lean expert) \\
    \textbf{TBD} & EPFL & 20 & 10 (Jul-Apr) & PhD Stud.\ (Lean expert) \\
    \textbf{TBD} & \textbf{TBD} & 40 & 2 (Jul-Sep) & Grad.\ Student (Proofreader) \\
    \hline\hline \rowcolor{lightgray}
    \multicolumn{3}{|l|}{\textbf{Total Hours}} & \multicolumn{2}{c|}{\textbf{3,320}} \\\hline
    \multicolumn{3}{|l|}{\textbf{Total Cost}} & \multicolumn{2}{c|}{\textbf{\TotalCost}} \\\hline
\end{tabular}%
}
\end{table}

@@

    \item Practitioners, on the other hand, often develop ad-hoc heuristics to make the Fiat--Shamir transformation work in practice. There are countless ad-hoc implementations in proof systems (e.g., \href{https://github.com/starkware-libs/stwo}{starkware-libs/stwo}, \href{https://github.com/Consensys/gnark-crypto}{Consensys/gnark-crypto}, \href{https://github.com/EspressoSystems/jellyfish}{EspressoSystems/jellyfish}, \href{https://github.com/arkworks-rs/sponge}{arkworks-rs/sponge}, etc.), several standards for specific applications (RFC 9591, RFC 9497, draft-irtf-cfrg-bbs-signatures, draft-google-cfrg-libzk, etc.), and a few generic frameworks (e.g., STROBE, SAFE API). None of these possess a formal security argument.
\end{itemize}

The main reason for this mismatch is that academic papers never describe how the Fiat--Shamir transformation can be built on top of fixed-size symmetric primitives, which is what symmetric cryptographers focus on. Because of the lack of a concrete specification, practitioners are left to fill in the gaps, leading to insecure implementations that often present the same class of bugs. Below, we highlight the main categories of well-known pitfalls that have led to insecure implementations:

\begin{itemize}
    \item \textbf{Incorrect serialization of the statement}. Failing to invoke the random oracle on the complete set of inputs, including the statement, session identifier, or protocol identifier, can enable proof forgery, replay, or cross-protocol attacks. Such mistakes have been uncovered in voting systems like \href{https://eprint.iacr.org/2016/771}{Helios}, \href{https://ieeexplore.ieee.org/document/9152765}{Scytl-SwissPost}, and \href{https://inria.hal.science/hal-02928953/file/final.pdf}{Belenios}, allowing for vote tampering, and more recently in some security audits:
    \begin{itemize}
    \item
\href{https://github.com/trailofbits/publications/blob/master/reviews/2023-04-scroll-zkEVM-wave1-securityreview.pdf}{\textsf{2023-04-scroll-zkEVM-wave1-securityreview}} (elliptic curve parameters omitted)
\item \href{https://github.com/Consensys/gnark/blob/master/audits/2024-09\%20-\%20Least\%20Authority\%20-\%20arithm\%20and\%20GKR.pdf}{\textsf{2024-09 - Least Authority - arithm and GKR}} (FFT domain generators omitted)
\item \href{https://leastauthority.com/wp-content/uploads/2024/11/Updated\_071124\_Polygon\_Plonky3\_Final\_Audit\_Report.pdf}{\textsf{Least Authority- Polygon Plonky3 Updated Final Audit Report}} (FRI configuration omitted)
\item \href{https://leastauthority.com/wp-content/uploads/2023/10/Safeheron_MPC_Algorithm_Updated_Final_Audit_Report_Least_Authority.pdf}{\textsf{Safeheron-MPC-Algorithm-Updated-Final-Audit-Report-Least-Authority}} (instance omitted)
    \end{itemize}
    \item \textbf{Erroneous implementation of the Fiat--Shamir transformation}. This includes implementations that perform the Fiat--Shamir transformation without placing all the prover messages in the Fiat--Shamir transformation. Such mistakes can lead to proof forgery, as previously seen in \href{https://www.youtube.com/watch?v=Sk-S8-n6Jo4}{the last challenge attack}, a widely used optimization for the \href{https://eprint.iacr.org/2025/565}{Gemini polynomial commitment scheme}, a recent \href{https://t.co/RFaHQOtAAq}{Solana bug},
    \href{https://github.com/trailofbits/publications/blob/master/reviews/zcoin-lelantus-summary}{zcoin-lelantus-summary.pdf}.

    \item \textbf{Incorrect mapping to and from the hash function domain}. This includes improper serialization of group elements, biased challenge generation, accepting multiple representatives of the same element, or failing to validate group membership. These issues are still affecting some commmon implementations such ask \href{https://github.com/arkworks-rs/crypto-primitives/blob/main/crypto-primitives/src/snark/constraints.rs#L163}{arkworks-rs} and need to be addressed.

    \item \textbf{Flawed random oracle methodology:} Some attacks exploit the gap between the theoretical random oracle model and its practical instantiation, undermining soundness. This includes the line of work \href{https://dl.acm.org/doi/10.5555/946243.946302}{started by Yael Kalai} which culminated in this year's \href{https://eprint.iacr.org/2025/118}{attack} to the Fiat-Shamir implementation used in \href{https://github.com/PolyhedraZK/Expander/}{PolyhedraZK/Expander}. This class of attacks is out of scope for this project proposal, but we will interact with other researchers working on this topic to ensure that our implementation can be adapted to potential mitigation strategies.
\end{itemize}

@@

\begin{table}[t]
    \def\arraystretch{1.2}%  1 is the default, change whatever you need
\centering
\caption{Tentative Gantt diagram illustrating the proposed timeline for the project.}
\resizebox{\textwidth}{!}{%
\begin{ganttchart}[y unit title=0.4cm,
    y unit chart=0.6cm,
    vgrid,hgrid,inline,
    title label anchor/.style={below=-1.6ex},
    title left shift=.05,
    title right shift=-.05,
    title height=1,
    progress label text={},
    x unit=2.5cm,
    bar height=0.7,
    group right shift=0,
    group top shift=.6,
    group height=.3]{1}{10}
    %labels
    \gantttitle{Fiat--Shamir Project (Jul--Apr 2026)}{10} \\
    \gantttitle{Jul}{1}
    \gantttitle{Aug}{1}
    \gantttitle{Sep}{1}
    \gantttitle{Oct}{1}
    \gantttitle{Nov}{1}
    \gantttitle{Dec}{1}
    \gantttitle{Jan}{1}
    \gantttitle{Feb}{1}
    \gantttitle{Mar}{1}
    \gantttitle{Apr}{1} \\
    % Work packages with step references
    % \ganttgroup[]{WP1}{1}{5} \\
    \ganttbar{WP1.1}{1}{5} \\
    \ganttbar{WP1.2}{2}{6} \\
    \ganttbar{WP1.3}{6}{7} \\
    % WP2
    % \ganttgroup[]{WP2}{3}{10} \\
    \ganttbar{WP2.1}{5}{7} \\
    \ganttbar{WP2.2}{7}{9} \\
    \ganttbar{WP2.3}{9}{10} \\
\end{ganttchart}
}
\end{table}

\paragraph{WP1: Formal Verification in Lean.}
This work package encompasses the formal verification of the Fiat--Shamir transformation of an interactive protocol. It will consist of the following steps.
\begin{enumerate}[label=\textbf{Step \arabic*:}, leftmargin=*, itemsep=0.5em]
    \item Prove Theorem~14.3.1, 14.4.1, 16.2.1 in the SNARG book, which allow us to bridge soundness, knowledge soundness, and zero-knowledge of the non-interactive protocol to the interactive protocol.
    \item Prove the Main Lemma (Lemma~5.1) from the paper. This step will carefully re-map any adversary for the designated Fiat--Shamir transformation to an adversary for the interactive protocol, in such a way that the resulting arguments strings produced have the same verification bit and the random oracle queries can be carefully re-mapped to random oracle queries.
    \item Prove soundness, knowledge soundness and zero-knowledge of the non-interactive protocol. This step will rely on the two previous steps, and will be done in a modular way, so that the transformation can be re-used by other protocols and other Fiat--Shamir transformations that guarantee properties similar to the main lemma of Step 2.
\end{enumerate}
\paragraph{WP2: Specification and Reference Implementation. }
This work package will focus on providing a specification of the Fiat--Shamir transformation, ideally reaching consensus with the broader cryptographic community.
\begin{enumerate}[label=\textbf{Step \arabic*:}, leftmargin=*, itemsep=0.5em]

@@

Our specification must appeal both traditional standards bodies and modern SNARK constructions. To address this, we will develop a modular architecture (e.g.\ abstracting away the hash function and the prover/verifier re-mappings) from the  to try to address inputs from the different communities involved in the process, letting different parties potentially edit different components. We will try to establish agreement and inter-operability with different projects active in the space.
Finally, we will document design decisions, develop test suites targeting known edge cases, and conduct coordinated security reviews with the cryptography engineering community before publication.

@@

\paragraph{Travel and Dissemination.} The travel and dissemination budget supports in-person collaboration and outreach activities. It includes a research visit for the project coordinators to meet in Lausanne at the COMPSEC lab of EPFL, enabling direct coordination on project milestones and addressing potential issues in the paper or implementation. Participation to academic conferences for the PI and the students is planned to build consensus within the cryptography and applied cryptography community around the Fiat--Shamir transformation. Attendance to these events will foster a broader community adoption. Finally, a visit to the Simons Institute in Berkeley will facilitate close collaboration with team members and other researchers participating in the \href{https://simons.berkeley.edu/workshops/cryptography-10-years-later-obfuscation-proof-systems-secure-computation-boot-campx}{Proofs Workshop} and working on the Lean implementation.

>>>>

\begin{table}[h]
    \caption{Summary table of persons involved in the project.}
    \centering
    \resizebox{\textwidth}{!}{%
\begin{tabular}{|c|l|c|c|l|}
    \hline
    Last, First name & Affiliation & Involvement (h/week) & Duration (months) & Role  \\\hline\hline
    \textbf{Orrù},  Michele & CNRS & 7 & 12 (Jul'25--Jun'26) & Project Coordinator \\
    \textbf{Chiesa}, Alessandro & EPFL & 5 & 12 (Jul'25--Jun'26) & Technical Advisor \\
    \textbf{TBD} & Ethereum Foundation & 20 & 12 (Jul'25--Jun'26) & PhD Student (Lean expert) \\
    \textbf{TBD} & Ethereum Foundation & 20 & 12 (Jul'25--Jun'26) & PhD Student (Lean expert) \\
    \textbf{TBD} & EPFL & 20 & 12 (Jul'25--Jun'26) & PhD Stud.\ (Lean expert) \\
    \textbf{TBD} & \textbf{TBD} & 40 & 2 (Jul-Sep'25) & Grad.\ Student (Proofreader) \\
    \hline\hline \rowcolor{lightgray}
    \multicolumn{3}{|l|}{\textbf{Total Hours}} & \multicolumn{2}{c|}{\textbf{4,928}} \\\hline
    \multicolumn{3}{|l|}{\textbf{Total Cost}} & \multicolumn{2}{c|}{\textbf{\TotalCost}} \\\hline
\end{tabular}%
}
\end{table}

@@

    \item Practitioners, on the other hand, often develop ad-hoc heuristics to make the Fiat--Shamir transformation work in practice. There are countless ad-hoc implementations in proof systems (e.g., \href{https://github.com/starkware-libs/stwo}{starkware-libs/stwo}, \href{https://github.com/Consensys/gnark-crypto}{Consensys/gnark-crypto}, \href{https://github.com/EspressoSystems/jellyfish}{EspressoSystems/jellyfish}, \href{https://github.com/arkworks-rs/sponge}{arkworks-rs/sponge}, etc.), several standards for specific applications (\href{https://datatracker.ietf.org/doc/rfc9591/}{RFC 9591}, \href{https://datatracker.ietf.org/doc/rfc9497/}{RFC 9497}, \href{https://datatracker.ietf.org/doc/draft-irtf-cfrg-bbs-signatures/}{draft-irtf-cfrg-bbs-signatures}, \href{https://datatracker.ietf.org/doc/draft-google-cfrg-libzk/}{draft-google-cfrg-libzk}, etc.), and a few generic frameworks (e.g., STROBE, SAFE API). None of these possess a formal security argument.
\end{itemize}

The main reason for this mismatch is that academic papers in the zero-knowledge space never formally describe how the Fiat--Shamir transformation can be built on top of fixed-length cryptographic hash functions, as designed by symmetric cryptographers. Because of the lack of a concrete specification, practitioners are left to fill in the gaps, leading to insecure implementations that often present the same class of bugs. Below, we highlight the main categories of well-known pitfalls that have led to insecure implementations:

\begin{itemize}
    \item \textbf{Incorrect serialization of the statement}. Failing to invoke the random oracle on the complete set of inputs, including the statement, session identifier, or protocol identifier, can enable proof forgery, replay, or cross-protocol attacks. Such mistakes have been uncovered in voting systems like \href{https://eprint.iacr.org/2016/771}{Helios}, \href{https://ieeexplore.ieee.org/document/9152765}{Scytl-SwissPost}, and \href{https://inria.hal.science/hal-02928953/file/final.pdf}{Belenios}, allowing for vote tampering, and more recently in some security audits:
    \begin{itemize}
    \item
\href{https://github.com/trailofbits/publications/blob/master/reviews/2023-04-scroll-zkEVM-wave1-securityreview.pdf}{\textsf{2023-04-scroll-zkEVM-wave1-securityreview}} (elliptic curve parameters omitted)
\item \href{https://github.com/Consensys/gnark/blob/master/audits/2024-09\%20-\%20Least\%20Authority\%20-\%20arithm\%20and\%20GKR.pdf}{\textsf{2024-09 - Least Authority - arithm and GKR}} (FFT domain generators omitted)
\item \href{https://leastauthority.com/wp-content/uploads/2024/11/Updated\_071124\_Polygon\_Plonky3\_Final\_Audit\_Report.pdf}{\textsf{Least Authority- Polygon Plonky3 Updated Final Audit Report}} (FRI configuration omitted)
\item \href{https://leastauthority.com/wp-content/uploads/2023/10/Safeheron_MPC_Algorithm_Updated_Final_Audit_Report_Least_Authority.pdf}{\textsf{Safeheron-MPC-Algorithm-Updated-Final-Audit-Report-Least-Authority}} (instance omitted)
    \end{itemize}
    \item \textbf{Erroneous implementation of the Fiat--Shamir transformation}. This includes implementations that perform the Fiat--Shamir transformation without placing all the prover messages in the Fiat--Shamir transformation. Such mistakes can lead to proof forgery, as previously seen in \href{https://www.youtube.com/watch?v=Sk-S8-n6Jo4}{the last challenge attack}, a widely used optimization for the \href{https://eprint.iacr.org/2025/565}{Gemini polynomial commitment scheme}, a recent \href{https://t.co/RFaHQOtAAq}{Solana bug},
    \href{https://github.com/trailofbits/publications/blob/master/reviews/zcoin-lelantus-summary}{zcoin-lelantus-summary}.

    \item \textbf{Incorrect mapping to and from the hash function domain}. This includes improper serialization of group elements, biased challenge generation, accepting multiple representatives of the same element, or failing to validate group membership. These issues are still affecting some common implementations such as \href{https://github.com/arkworks-rs/crypto-primitives/blob/main/crypto-primitives/src/snark/constraints.rs#L163}{arkworks-rs} and need to be addressed.

    \item \textbf{Flawed random oracle methodology:} Some attacks exploit the gap between the theoretical random oracle model and its practical instantiation, undermining soundness. This includes the line of work \href{https://dl.acm.org/doi/10.5555/946243.946302}{started by Yael Kalai}, which culminated in this year's \href{https://eprint.iacr.org/2025/118}{attack} on the Fiat--Shamir implementation used in \href{https://github.com/PolyhedraZK/Expander/}{PolyhedraZK/Expander}. This class of attacks is out of scope for this project proposal. We will however interact with other researchers working on this topic so that our implementation can be adapted to potential mitigation strategies.
\end{itemize}

@@

\begin{table}[t]
    \def\arraystretch{1.2}%  1 is the default, change whatever you need
\centering
\caption{Tentative Gantt diagram illustrating the proposed timeline for the project.}
\resizebox{\textwidth}{!}{%
\begin{ganttchart}[y unit title=0.4cm,
    y unit chart=0.6cm,
    vgrid,hgrid,inline,
    title label anchor/.style={below=-1.6ex},
    title left shift=.05,
    title right shift=-.05,
    title height=1,
    progress label text={},
    x unit=2.5cm,
    bar height=0.7,
    group right shift=0,
    group top shift=.6,
    group height=.3]{1}{12}
    %labels
    \gantttitle{Fiat--Shamir Project (Jul'25--Jun'26)}{12} \\
    \gantttitle{Jul'25}{1}
    \gantttitle{Aug}{1}
    \gantttitle{Sep}{1}
    \gantttitle{Oct}{1}
    \gantttitle{Nov}{1}
    \gantttitle{Dec}{1}
    \gantttitle{Jan'26}{1}
    \gantttitle{Feb}{1}
    \gantttitle{Mar}{1}
    \gantttitle{Apr}{1}
    \gantttitle{May}{1}
    \gantttitle{Jun'26}{1} \\
    % Work packages with step references
    % \ganttgroup[]{WP1}{1}{12} \\
    \ganttbar{WP1.1}{1}{6} \\
    \ganttbar{WP1.2}{5}{9} \\
    \ganttbar{WP1.3}{9}{10} \\
    % WP2
    % \ganttgroup[]{WP2}{7}{12} \\
    \ganttbar{WP2.1}{7}{10} \\
    \ganttbar{WP2.2}{10}{11} \\
    \ganttbar{WP2.3}{11}{12} \\
\end{ganttchart}
}
\end{table}

\paragraph{WP1: Formal Verification in Lean.}
This work package encompasses the formal verification of the Fiat--Shamir transformation of an interactive protocol. It will consist of the following steps.
\begin{enumerate}[label=\textbf{Step \arabic*:}, leftmargin=*, itemsep=0.5em]
    \item Prove Theorems~14.3.1, 14.4.1, 16.2.1 in the SNARG book, which allow us to bridge soundness, knowledge soundness, and zero-knowledge of the non-interactive protocol to the interactive protocol.
    \item Prove the Main Lemma (Lemma~5.1) from the paper. This step will carefully re-map any adversary for the designated Fiat--Shamir transformation to an adversary for the interactive protocol, in such a way that the resulting arguments strings, and the random oracle queries, can be re-mapped to random oracle queries.
    \item Prove soundness, knowledge soundness and zero-knowledge of the non-interactive protocol. This step will rely on the two previous steps, and will be done in a modular way, so that the transformation can be re-used by other protocols and other Fiat--Shamir transformations that guarantee properties similar to the main lemma of Step 2.
\end{enumerate}
\paragraph{WP2: Specification and Reference Implementation. }
This work package will focus on providing a specification of the Fiat--Shamir transformation, ideally reaching consensus with the broader cryptographic community.
\begin{enumerate}[label=\textbf{Step \arabic*:}, leftmargin=*, itemsep=0.5em]

@@

Our specification must appeal to both traditional standards bodies and modern SNARK constructions. To address this, we will develop a modular architecture (e.g., abstracting away the hash function and the prover/verifier re-mappings) to address input from the different communities involved in the process, allowing different parties to edit different components. We will seek to agreement and interoperability with other active projects in the space.
Finally, we will document design decisions, develop test suites targeting known edge cases, and conduct security reviews with the cryptographic engineering community before publication.

@@

\paragraph{Travel and Dissemination.} The travel and dissemination budget supports in-person collaboration and outreach activities. It includes a research visit for the project coordinators to meet in Lausanne at the COMPSEC lab of EPFL, enabling direct coordination on project milestones and addressing potential issues in the paper or implementation. Participation in academic conferences for the PI and the students is planned to build consensus within the cryptography and applied cryptography community around the Fiat--Shamir transformation. Attendance at these events will foster broader community adoption. Finally, a visit to the Simons Institute in Berkeley will facilitate close collaboration with team members and other researchers participating in the \href{https://simons.berkeley.edu/workshops/cryptography-10-years-later-obfuscation-proof-systems-secure-computation-boot-campx}{Proofs Workshop} and working on the Lean implementation.
